---
title: "Class 18"
format: docx
execute:
  error: TRUE
---

# Plan
* Hypothesis testing + p-values (lecture)
* Visualizing hypothesis testing in R
* Group work review
* Quick practice (conceptual)
* Regression tables (lecture)
* Group Project Results
* Review session

# Lecture: Hypothesis testing 

# Visualizing the null hypothesis 

```{r}
library(tidyverse)
# Do not need to know the plotting code -- just a demonstration!

resume <- read.csv("./class_data/resume.csv")
chechen <- read.csv("./class_data/chechen.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/resume.csv","resume.csv")
#resume <- read.csv("resume.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/chechen.csv","chechen.csv")
#chechen <- read.csv("chechen.csv")

###### REJECTING THE NULL

resume_b <- resume |> filter(race=="black")
resume_w <- resume |> filter(race=="white")
t.test(resume_b$call,resume_w$call,conf.level=.95)

# or
out <- t.test(call ~ race, data=resume)
out
t <- out$statistic

#### Plot
n  <- nrow(resume)
conflevel = .95
x <- seq(-8, 8, length=100)
plot(x,dt(x,n-1),lty=1,type="l",xlab="Standard Deviations of Sampling Distribution (Centered on 0)")
abline(v=qt((1-conflevel)/2,n-1)*-1,col="red")
abline(v=qt((1-conflevel)/2,n-1),col="red")
abline(v=t,col="blue")

result <- lm(call ~ race, data=resume)
summary(result)
confint(result)

```

```{r}
###### FAIL TO REJECT THE NULL

out <- t.test(postattack ~ fire, data=chechen)
out
t <- out$statistic

#### Plot
n  <- nrow(chechen)
conflevel = .95
x <- seq(-8, 8, length=100)
plot(x,dt(x,n-1),lty=1,type="l",xlab="Standard Deviations of Sampling Distribution (Centered on 0)")
abline(v=qt((1-conflevel)/2,n-1)*-1,col="red")
abline(v=qt((1-conflevel)/2,n-1),col="red")
abline(v=t,col="blue")


result <- lm(postattack ~ fire, data=chechen)
summary(result)
confint(result)

```


# Lecture: p-values

# Review of group work

```{r}
library(tidyverse)

boston <- read.csv("./problem_sets/data/boston.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/problem_sets/data/boston.csv","boston.csv")
#boston <- read.csv("boston.csv")

```

Evaluate whether individuals in the treated and control group are comparable with respect to background characteristics (i.e., the difference in means is 0), using a 95% confidence level. The background characteristics are: age, income, male, white, college, usborn, partisanship.

For any characteristic, do we reject the possibility that the difference between the treated and control averages is 0? What does this imply about the quality of the randomization?

```{r}
# Different ways to do this:

# Age
summary(lm(age~treated,data=boston))
t.test(age~treated,data=boston)

# If we have multiple groups:
boston_t <- boston |> filter(treated==1)
boston_c <- boston |> filter(treated==0)
t.test(boston_t$age,boston_c$age,data=boston)


# Usborn
summary(lm(usborn~treated,data=boston))
t.test(usborn~treated,data=boston)
t.test(boston_t$usborn,boston_c$usborn,data=boston)

```

Re-evaluate the Kenyan savings experiments:

```{r}
library(tidyverse)
kenya <- read.csv("./class_data/rosca2.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/rosca2.csv","rosca2.csv")
#kenya <- read.csv("rosca2.csv")
kenya <- subset(kenya,has_followup2 == 1)
```

Variables:
fol2_amtinvest  Amount invested in Kenyan shillings, in Wave 2 (outcome)
encouragement   Binary variable indicating whether individuals are in the control group (1) or not (0)
safe_box        Binary variable indicating whether individuals are in the safe deposit box treatment group (1) or not (0)
locked_box      Binary variable indicating whether individuals are in the locked box treatment group (1) or not (0)
bg_b1_age       age of participant
bg_female       gender of participant
bg_married      marital status of participant

Calculate the t-statistic (at the 95% level) for the null hypothesis that the difference in fol2_amtinvest between the locked box group and the control group is 0. Repeat for the safe box group. 
 
```{r}
# Note: since we have multiple treatment conditions, it is better to filter and then use t.test

lb <- kenya |> filter(locked_box == 1)
sb <- kenya |> filter(safe_box == 1)
c <- kenya |> filter(encouragement == 1)

t.test(lb$fol2_amtinvest, c$fol2_amtinvest)
# fail to reject (statistically insignificant)

t.test(sb$fol2_amtinvest, c$fol2_amtinvest)
# reject (statistically significant)

# Alternative using lm(). The t values will differ here due to a difference in degrees of freedom; conclusions remain unchanged
result <-lm(fol2_amtinvest ~ locked_box + safe_box,data=kenya)
summary(result)

```

Now evaluate balance between the treated and control conditions with respect to age, gender, and marital status, at the 95% confidence level. Note: since there are two treatment conditions, you will have to evaluate the balance of each treatment relative to control. The easiest way to do this is with a t-test, but you could also use regression if you created datasets that exclude the other treatment group.

For any characteristic, at 95% confidence do we reject the possibility that the difference between the treated and control averages is 0? What does this imply about the quality of the randomization?

```{r}
# Locked box
t.test(lb$bg_b1_age, c$bg_b1_age)
t.test(lb$bg_married, c$bg_married)
t.test(lb$bg_female, c$bg_female)
# Fail to reject at 95% level for all

# Safe box
t.test(sb$bg_b1_age, c$bg_b1_age)
t.test(sb$bg_married, c$bg_married)
t.test(sb$bg_female, c$bg_female)

# Fail to reject at 95% level for all; but would reject at
# 90% confidence for age

# Overall, the groups are well balanced (good randomization)
```


# Quick Practice

1. You are evaluating an RCT and examining whether the treated and control groups are similar with respect to background characteristics. You conduct
a hypothesis test where you hypothesize that the difference in the share of females in each group is 0. Evaluating the hypothesis test, you see a p-value of 0.4

Do you reject or fail to reject the null hypothesis?

What does this imply about the quality of randomization for this variable?

2. Your friend Brunhilda tests another variable (age), evaluating the null hypothesis that the difference in average age between the treated and control groups is 0. She receives a t-statistic of 4. The sample size for the RCT is 2000. 

Do you reject or fail to reject the null hypothesis?

What does this imply about the quality of randomization for this variable?

3. Now you examine the ATE. The 95% confidence interval for the ATE is -0.04 to -0.01. Do you reject or fail to reject the null hypothesis of no effect, at the 95% level?

4. You examine a different study. The p-value on the ATE is .02. Is the finding statistically significant at the 95% level?

# Lecture: Regression tables

# Group Project Results

# Final logistics

In-person exam: 1 hr 30 mins (Monday 11:30am, this room)
Take-home exam: 45 mins (posted by Fri)

# Review session (conceptual problems)

### Scratchpad for conceptual problems

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
