---
title: "Class 17"
format: docx
execute:
  error: TRUE
---

# Plan
* Quick review: confidence intervals
* Lecture: t-distribution
* t-distribution in R
* Lecture: hypothesis testing
* Group work

# Quick review:

ATE / Confidence Intervals with Regression

```{r}
gotv <- read.csv("./class_data/social2.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/social2.csv","social2.csv")
#gotv<- read.csv("social2.csv")

# ATE using regression
result <- lm(primary2008~ messages,data=gotv)
summary(result)

```

```{r}
gotv <- gotv |> mutate(
  civic=ifelse(messages=="Civic Duty",1,0),
  hawthorne=ifelse(messages=="Hawthorne",1,0),
  neighbors=ifelse(messages=="Neighbors",1,0)
)

# ATE using regression

result <- lm(primary2008~ civic+hawthorne+neighbors,data=gotv) # leave out control group
summary(result)


```


#  Lecture: t-distribution

Compare t-distribution to the normal distribution 

```{r}
# Do not need to know this code
n <- samplesize <- 4
x <- seq(-4, 4, length=100)
hx <- dnorm(x)
plot(x,hx,lty=1,type="l",col="blue")
lines(x, dt(x,n-1), lwd=2, col="red")
```

For smaller sample sizes, we can use _qt()_ to obtain critical values and confidence intervals. This is the equivalent of _qnorm_ for the t-distribution. qt() requires us to state the degrees of freedom (n-1).

Let's obtain critical values for a 95% confidence interval, with a sample size of 10

```{r}
#qt(,df=)
```

We can use qt() all the time to be conservative, but if the sample size is large the critical value will be similar to qnorm(). Let's use qt() for the resume example, constructing a 90% confidence interval

```{r}

```

# Lecture: Hypothesis testing


Our t-statistic is 1.9. Do we reject the null hypothesis of 0 effect, or fail to reject?

```{r}
confidencelevel <- (1-.9)/2
df <- 
qt(confidencelevel,df)

```


# t-tests in R

We can generate t-statistics manually. However, the _t.test_ command does this automatically:

Let's test whether the callback rate for black resumes is different from 0, at the 99% level.

## One sample t-test (for a single distribution):

By default t.test evaluates the null hypothesis of 0:

```{r}
library(tidyverse)
resume <- read.csv("./class_data/resume.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/main/class_data/resume.csv","resume.csv")
#resume <- read.csv("resume.csv")
resume <- resume %>% mutate(female=ifelse(sex=="female",1,0))

resume_b <- resume |> filter()

t.test(,conf.level=.99)

```

We can also test whether it is different from another value (such as 2%), using the ,mu option:

```{r}
t.test(,mu=.02,conf.level=.99)

```

## Two sample t-test  (t-test for difference in means):

Let's test whether the difference in callback rates for black and white resumes is different from 0, at the 95% level. To do a two sample t-test, we put in each variable, separated by commas

```{r}
t.test(resume_b$call,resume_w$call,conf.level=.95)

```

We can also write it like this:

```{r}
t.test(~,conf.level=.95)

```

We can also use t-tests to check for accurate randomization. Let's check that gender was independently randomized between the white and black resumes, at the 95% level. If randomization is accurate, we should fail to reject the null hypothesis that the difference in means is 0.

```{r}


```

We can use regression to obtain t-statistics. But note that regression *always* tests the null hypothesis that the true coefficient is 0. 

```{r}

result <- lm(call~race,data=resume)
summary(result)

```

When evaluating randomization quality with regression, the dependent variable should be the variable we expect to be randomized, and the independent variable should be the treatment indicator. 

```{r}

result <- lm(,data=resume)
summary(result)

```


# Group Work: Practice with Hypothesis Testing 

Choose a study below, and (a) determine whether we can reject the null hypothesis that the average treatment effect is 0, and (b) evaluate the quality of the randomization by testing whether the difference between the treated and control groups for a given characteristic is different from 0.

##A

Re-evaluate the Boston immigration experiment (from problem set 1)

| Name                  | Description                                                     |
|-----------------------|-----------------------------------------------------------------|
| age                   | Age of individual at time of experiment                                   
| treatment             | Whether an individual was exposed to immigration and is in the treatment group (1) or (0) 
| income                | Income group in dollars (not exact income)
| male                  | Whether individual indicated they were male (1) or not (0)
| white                 | Indicator variable for whether individual identifies as white (1) or not (0)
| college               | Indicator variable for whether individual attended college (1) or not (0)
| usborn                | Indicator variable for whether individual is born in the US (1) or not (0)
| partisanship          | Self-reported political views, ranging from (1) Very Liberal to (5) Very Conservative
| immig_after           | The outcome. Policy opinion on question about increasing the number of immigrants allowed in the country, on a scale 
                        | from Oppose Increase (1) to Not Sure (3) to Support Increase (5). 

```{r}
library(tidyverse)

boston <- read.csv("./problem_sets/data/boston.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/problem_sets/data/boston.csv","boston.csv")
#boston <- read.csv("boston.csv")

```

Recall that the outcome in this study is attitudes toward immigration (immig_after), while the treatment variable indicates whether someone encountered Hispanic travelers on their morning commute. 

Calculate the average treatment effect, and obtain a 95% confidence interval for the ATE. Obtain a t-statistic and then evaluate the null hypothesis that the true effect of immigration exposure on attitudes is 0. 

```{r}


```

*Answer here*

Evaluate whether individuals in the treated and control group are comparable with respect to background characteristics (i.e., the difference in means is 0), using a 95% confidence level. The background characteristics are: age, income, male, white, college, usborn, partisanship.

For any characteristic, do we reject the possibility that the difference between the treated and control averages is 0? What does this imply about the quality of the randomization?

```{r}

```

*Answer here*


##B

Re-evaluate the Kenyan savings experiments.

```{r}
library(tidyverse)
kenya <- read.csv("./class_data/rosca2.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/rosca2.csv","rosca2.csv")
#kenya <- read.csv("rosca2.csv")
kenya <- subset(kenya,has_followup2 == 1)
```

Variables:
fol2_amtinvest  Amount invested in Kenyan shillings, in Wave 2 (outcome)
encouragement   Binary variable indicating whether individuals are in the control group (1) or not (0)
safe_box        Binary variable indicating whether individuals are in the safe deposit box treatment group (1) or not (0)
locked_box      Binary variable indicating whether individuals are in the locked box treatment group (1) or not (0)
bg_b1_age       age of participant
bg_female       gender of participant
bg_married      marital status of participant

Calculate the t-statistic (at the 95% level) for the null hypothesis that the difference in fol2_amtinvest between the locked box group and the control group is 0. Repeat for the safe box group. 
 
```{r}

```

*Answer here*

Now evaluate balance between the treated and control conditions with respect to age, gender, and marital status, at the 95% confidence level. Note: since there are two treatment conditions, you will have to evaluate the balance of each treatment relative to control. The easiest way to do this is with a t-test, but you could also use regression if you created datasets that exclude the other treatment group.

For any characteristic, at 95% confidence do we reject the possibility that the difference between the treated and control averages is 0? What does this imply about the quality of the randomization?

```{r}

```

*Answer here*



# If time: p-values


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
