---
title: "Class 9"
format: docx
execute:
  error: TRUE
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan

* Announcements
* Quiz review
* Correlation
* Standard Deviation 
* Sampling Error
* Sensitive Survey Questions

# Quiz Review

# Correlation

```{r}
library(tidyverse)
arson <- read.csv("./class_data/arson.csv")

#download.file("https://github.com/resonance1/govt10-f25/raw/refs/heads/main/class_data/arson.csv","arson.csv")
#arson<- read.csv("arson.csv")

```

```{r}
y <- arson$attack_rate
x <- arson$foreign
  
plot(x,y)
lines(lowess(x,y))

cor(x,y,use="complete.obs")

```

## Correlation tables

```{r}

```

# Short lecture: Standard Deviation

## Standard deviation in R

To calculate the standard deviation in R, we use the sd() function

```{r}

library(tidyverse)

hsavings <- read.csv("./class_data/rosca.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/rosca.csv","rosca.csv")
#hsavings <- read.csv("rosca.csv")

control <- hsavings |> filter(tgroup=="Control")

```

Let's visualize this with a histogram:

```{r}


```

We can use standard deviation to contextualize an ATE, by comparing the effect size to the standard deviation of the outcome in the control group:

```{r}
# ATE

hsavings |> group_by(tgroup) |> summarize(mean(fol2_amtinvest,na.rm=T))

408.2150	- 257.8333	

```

```{r}
# Visual demonstration - don't worry about the code

ate <- 150.3817

control <- hsavings |> filter(tgroup=="Control")
mean <- mean(control$fol2_amtinvest,na.rm=T)
sd <- sd(control$fol2_amtinvest,na.rm=T)
effect_1sd <- mean + sd
effect_2sd <- mean + 2*sd

hist(control$fol2_amtinvest,breaks=20)
abline(v=,col="")


```


# Short Lecture: Measurement and Sampling Error


## Sampling Error

Supports Trump    = 1
Does not support Trump = 0 

The support for Trump in this particular state is 40%:

```{r}
state <- c(rep(1,40000),rep(0,60000))

mean(state)

```

To take a sample, we use the _sample()_ function. The format is:

  sample(whatwewanttosample,sizeofsample,replace=F)

Let's draw a sample of size 10, and examine the mean.

```{r}
sample <- sample(state,10,replace=F)
mean(sample)

```

Now let's draw new samples of 100 and 500, and compare the mean.

```{r}
sample <- sample(state,,replace=F)
mean(sample)

```

Now let's draw a 1000 person sample. What is the expected error, using the equation covered in class?

```{r}

sample <- sample(state,1000,replace=F)

```

The error tells us how far off our sample mean will be from the population mean, on average. Technically, it gives us the standard deviation of the hypothetical sampling distribution: the standard deviation of all possible answers we could have received.

To test whether our formula for the expected error works, let's draw 100 samples, each of size 1000. Then we'll plot the average value from each sample. A distribution of sample means is known as a *sampling distribution* 

```{r}

# You do not need to know this code
n<-1000
samples <- 100
out <- c()
for (i in 1:samples){
  out <- c(out,mean(sample(state,n,replace=FALSE)))
}

hist(out)
sd(out)
mean(out)

```

### Practice question:

A polling company randomly sampled 500 Americans to ask them if they approved supplying arms to Ukraine. On a scale of 1 to 5, where 1 meant "strongly oppose" and 5 meant "strongly support", the average response was 4.12. The standard deviation was 0.25. What was the expected error of this poll?

```{r}

```



# Short Lecture: Other Sources of Polling Error

Let's load a survey of Afghan citizens. 

```{r}
afghan <- read.csv("./class_data/afghan.csv") %>% filter(list.group!="ISAF")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/afghan.csv","afghan.csv")
#afghan<- read.csv("afghan.csv") %>% filter(list.group!="ISAF")

```

One of the most common items individuals refuse to respond to in surveys are questions about income. How many fall into this category in the afghan survey?

To determine this, we can use is.na() on the _income_ variable.

```{r}

```

How does non-response vary by other characteristics? Let's create a variable that measures income non-response. Then let's examine how it varies as a function of employment

```{r}


```

# Short Lecture: Sensitive Questions

## Group Work: List Experiment (if time)

Now let’s assess the list experiment. The relevant variables are:

list.group      Whether respondents saw no sensitive items (Control), or a sensitive item stating that they support the Taliban
list.response   Number of items on the list respondents agreed with

1. What is the non-response rate to the list experiment?

```{r}

```

2. What percentage of respondents supported the Taliban?

```{r}


```





## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand (leave blank if nothing)?
