---
title: "Class 14"
format: docx
execute:
  error: TRUE
---

# Plan
* Quick Review on Creating Predictive Models
* Importing new data
* Regression and ATE (lecture)
* Group work
* Limits of multivariate models 
* Interaction effects (lecture)
* Interaction effects in R

# Quick Review on Creating Predictive Models 

See template.qmd

# Importing new data

```{r}
# data <- read.csv("")

# Merge to training dataset, by state
```

# Brief Lecture: Regression and Causation

# Group work

Let's examine a simple case first, with 1 treatment. 

We'll load the resume racial discrimination study, from the very beginning of the course:

```{r}
resume <- read.csv("./class_data/resume.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/resume.csv","resume.csv")
#resume<- read.csv("resume.csv")
```

First, calculate the ATE for the experiment using group_by |> summarize().

The outcome variable is *call* (1 or 0), and the treatment variable is *race*. Let's assume that the treated group is when race=="white" and the control is when race=="black"

```{r}

```

Now do it with a regression. For this, assign *call* to the y and *race* to the x, using the lm() function.

```{r}

```

Now let's examine the Kenyan savings experiment again:

```{r}
kenya <- read.csv("./class_data/rosca2.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/rosca2.csv","kenya.csv")
#kenya<- read.csv("kenya.csv")

kenya <- subset(kenya,has_followup2 == 1)
```

fol2_amtinvest            The outcome (investment in Kenyan shillings)
encouragement             1 if the person is in the control group, 0 otherwise
locked_box                1 if the person is in the locked box group, 0 otherwise
safe_box                  1 if the person is in the safe_box group, 0 otherwise
bg_b1_age                 Age of participant
bg_married                Whether participant is married (1) or not (0)
bg_female                 Whether participant is female (1) or not (0)

Calculate the ATE of the locked_box and safe_box groups, using a single multivariate regression model. You'll have to omit the control group for this to work. 

```{r}

```

Now add all the background characteristics of participants to the model. Do the ATEs change? What might this tell you about the quality of the randomization in this study? Interpret the coefficient on Locked Box.

```{r}


```

# As a class: Limits of multivariate models

*Degrees of freedom:*
This corresponds to our number of observations (n), minus 1, minus the number of independent variables in our model. Degrees of freedom must always be positive to return coefficients. If degrees of freedom is too small, we also become more uncertain about whether our coefficients are accurate.

*Collinearity:*
If independent variables are perfectly correlated, we cannot estimate both terms. Very high correlation is also problematic for accurate coefficients

```{r}

# Number of observations

kenya_subset <- kenya %>% slice_sample(5)
result <- lm(fol2_amtinvest ~ safe_box + locked_box + bg_b1_age + bg_married + bg_female, data = kenya_subset)
summary(result)

# Collinearity

kenya <- kenya %>% mutate(bg_unmarried = ifelse(bg_married==0,1,0))
kenya$bg_b1_age_18 = kenya$bg_b1_age - 18

result <- lm(fol2_amtinvest ~ safe_box + locked_box + bg_b1_age + bg_married  + bg_female, data = kenya)
summary(result)
```

# Interaction effects (lecture)

# Interaction effects in R

Heterogeneity by gender

```{r}
gotv <- read.csv("./class_data/social2.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/social2.csv","social2.csv")
#gotv<- read.csv("social2.csv")

gotv <- gotv |> filter(messages=="Control" | messages == "Neighbors")

result <- lm(primary2008 ~ messages, data=gotv)
result

```

What if we use interaction effects with continuous variables?

Household Size (hhsize)

```{r}

```

Visualization

```{r}
#install.packages("sjPlot")
#library(sjPlot)


plot_model(result,type="pred",terms=c("hhsize","messages"))

```

# Only if time: Group Work - Interaction Effects

Let's look at a simplified version of the Kenyan savings example

```{r}
# Create a subset of the dataframe called kenya2 with just the locked box treatment
kenya2 <- kenya |> filter(safe_box!=1) |> select(-safe_box,-has_followup2,-X)

```

Separately explore interactions with marital status, gender, and age. Choose the regression with the strongest interaction effect and interpret the beta coefficients

```{r}

result <- lm(fol2_amtinvest ~ locked_box, data=kenya2)
result

```



## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
