---
title: "Class 10"
format: docx
execute:
  error: TRUE
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan

* Quick review
* Sensitive survey questions
* Prediction

# Review: Standard deviation vs standard error

*Standard deviation* is a summary statistic that we apply to distributions. It indicates how far, on average, observations are from the mean.

*Standard error* tells us how different our **sample** mean is likely to be from the true **population** mean. In other words, what is the expected error of our estimate? Technically, the standard error is the standard deviation of the (hypothetical) sampling distribution.

# Recap: Sampling Distribution

Create a population:

```{r}
state <- c(rep(1,40000),rep(0,60000))
```

Sample from the population, with a sample size (n) large enough for the Law of Large Numbers:

```{r}
sample <- sample(state,1000,replace=F)
```

The error tells us how far off our sample mean will be from the population mean, on average

```{r}
n <- 1000
sd <- sd(sample)

sd/sqrt(n)

```

This is the same as the standard deviation of the hypothetical sampling distribution: the standard deviation of all possible answers we could have received

```{r}
# You do not need to know this code. This repeats the sampling procedure we did above 10,000 times
n<- 1000
samples <- 10000
out <- c()
for (i in 1:samples){
  out <- c(out,mean(sample(state,n,replace=FALSE)))
}

hist(out,breaks=100)
mean(out)
sd(out)
```


# Short Lecture: Sensitive Questions

## List Experiment

```{r}
library(tidyverse)
afghan <- read.csv("./class_data/afghan.csv") %>% filter(list.group!="ISAF")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/afghan.csv","afghan.csv")
#afghan<- read.csv("afghan.csv") %>% filter(list.group!="ISAF")

```

Now let’s assess the list experiment. The relevant variables are:

list.group      Whether respondents saw no sensitive items (Control), or a sensitive item stating that they support the Taliban
list.response   Number of items on the list respondents agreed with

1. What is the non-response rate to the list experiment? Note: you'll have to use is.na() for this. Review from class9.qmd if needed

```{r}

```

2. What percentage of respondents supported the Taliban?

```{r}


```

# Short Lecture: Prediction

## 2008 Presidential Election

Load the dataset:

```{r}
library(tidyverse)
polls08 <- read.csv("./class_data/polls08.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/polls08.csv","polls08.csv")
#polls08 <- read.csv("polls08.csv")

```

Let's create a new variable called _poll.margin_ that measures the difference in polling support between Obama and McCain:

```{r}

polls08 <- polls08 |> mutate(poll.margin=Obama-McCain)

```

Let’s examine polls within the state of California by filtering

```{r}

polls08 |> filter()

```

We can select the first observation using slice(1)

```{r}


```

We can do this for all states at once by returning to the main dataset and using |> group_by _before_ slicing. 

```{r}


```

## Merging datasets

What if we want to compare the accuracy of our poll results to the actual results? To do that, we will need to merge two datasets together.

Load the actual results: 

```{r}
pres08 <- read.csv("./class_data/pres08.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/main/class_data/pres08.csv","pres08.csv")
#pres08 <- read.csv("pres08.csv")

```

Let's create a margin variable in this new dataset that mirrors our prior format (Obama vs Mccain)

```{r}
pres08 <- pres08 |> mutate(act.margin=Obama-McCain)

```

To merge datasets, we need to find a _key_ that's held in common between the two datasets.

```{r}

```

We can merge using the following syntax:

dataset1 |> left_join(dataset2,by="key")

```{r}

pres08 |> left_join()


```

This can result in some duplication of columns. To fix this, we can trim a dataset before merging using select():

```{r}


```

## Error

Evaluating Prediction error: Actual - Predicted

```{r}
# Create error variable
master <- master |> mutate()

```

Root mean squared error (RMSE):

sqrt(mean(error^2))

```{r}

```


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand (leave blank if nothing)?

