---
title: "Class 1"
format: docx
execute:
  error: TRUE
---

# Class Plan

- Class information 
- Project management with Rstudio / git 
- Markdown
- R tutorial: Basic operations
- R tutorial: Variables
- R tutorial: Data frames

# Rstudio intro

# R notebooks

R Notebooks allow us to write text, write code and generate results within a single file. In this class all in-class assignments and homework assignments will be submitted using Markdown Notebooks (qmd).

Markdown allows us to include code and format text. Markdown can be compiled directly into a Word file or an HTML file.

# Markdown

For this class we only care about some basic formatting:

**Bolding text** *Italicizing text*

The **quick** brown fox jumps over the *startled* gopher

# Creating a larger header

## Creating a second level header

## How to write code within markdown

When you see a white background you are writing text. You can use Markdown here. However we will also be interested in writing code to explore and summarize data. Code is placed in a *chunk.* Chunks start and end with three grave accents (the same key as the tilde): chunks are interactive and allow to you run code. When you execute code within the notebook, the results appear beneath the code. To run a chunk, press the *Run* button (play button) within the chunk.

Multiply 5 and 10:

```{r}

```


```{r}

```


```{r}
# NOTE: we cannot use markdown in code chunks.  Here # means that we are creating a comment (notes that are not executable code).  We are not creating a header.

```

Multiply 4 and 6, then square it

```{r}

```

R also contains *functions*. For instance, the function sqrt() calculates a square root:

```{r}

```

# Group work: Exercise 1

Expanding this notebook, complete the following tasks:

1)  Create two new code chunks below this text (using the green 'C' button in the top right).

2)  In the first chunk, find the sum of 75, 83, 93, 34, 75 and 81 using math operators (+). Then try to find the sum using the function *sum()*. You can look up how to use the function *sum()* by typing ?sum in the console (bottom of your screen). Or ask Claude/GPT to show you how it works.

3)  In the second chunk divide 75 by 5, then add 43 and then multiply 6 (use parentheses)

# On your own: Exercise 2

Using markdown, answer the following questions directly in this notebook:

1)  Where are you from?
2)  What is your class year?
3)  What is your major/planned major?
4)  Have you taken a statistics or a programming class before?
5)  What is your favorite book or movie?

When you've finished this, close your computer so I know when you're ready for the next part of the class.

# As a class: Variables

R can save information as a variable using the following syntax: *variable name <- value*

Let's create a variable named my_variable, with the value of 13.

```{r}

```

We can also use '=' to create a variable

```{r}


```

Get the value of the variable *my_variable* by typing the variable name

```{r}

```

There are some limits to variable names. Don't use spaces, or (most) punctuation. Dots, dashes, underscores are fine.

```{r}


```

We can input variable names into functions. Let's take the natural log of *my_variable*

Note: There are many helpful lists of functions online: <https://www.statmethods.net/management/functions.html>

```{r}


```

Let's change the value of *my_variable* by overwriting it

```{r}


```

Now let's copy the variable *my_variable* to a new variable called *my_variable2*

```{r}


```

We can perform mathematical operations directly on variables

```{r}


```


## String (text) variables

We can also store text within variables by wrapping the text in double quotation marks (""). Let's create a new variable called my_text that contains the words "gov 10"

```{r}


```

We can't perform mathematical operations on text

```{r}


```

# Data frames

Data frames are similar to a spreadsheet, and contain multiple observations (rows) and variables (columns). We'll also call them "data sets" in this course

## Loading data frames

For this class, all data are stored in .csv format. Let's load turnout.csv. There are two ways we can do this.

A)  Sync git, then read in the file:

```{r}
turnout <- read.csv("class_data/turnout.csv")
```

B)  Download the file from the internet into your current directory and load it:

```{r}
#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/turnout.csv","turnout.csv")
#turnout <- read.csv("turnout.csv")
```

## Data frames and variables

We can see the variable names in our data frame using *names()*

```{r}

```

We can pull out the first few rows using *head()*

```{r}

```

We can select columns in a dataframe by name using the *$* operator. When using the *$ operator we place the name of the data frame to the left and the name of the variable to the right.

Let's pull out the column *year*

```{r}


```

We can run functions on columns:

```{r}

```

To create a new column in our data frame, we can also use the *$* operator. If we pass it a single value, that value will be placed within every row.

```{r}

```

We can also create variables that are combinations of existing variables in the dataset. Let's create a variable called **VEP_turnout** that shows the actual turnout rate.

```{r}

```

# Group work: Social desirability bias

Surveys are frequently used to measure political behavior such as voter turnout, but some researchers are concerned about the accuracy of self-reporting. In particular, they worry about possible *social desirability bias*. In this phenomenon, respondents lie on surveys out of a desire to provide the 'socially correct' answer. In this case, those who did not vote in an election may report that they did.

To study this issue, we can look at data from the ANES. The ANES is a nation-wide survey that has been conducted for every election since 1948. The ANES conducts face-to-face interviews with a nationally representative sample of adults.

1. The variable we created, VEP_turnout, measures the actual turnout rate recorded by election authorities. Using this variable, what is the average turnout rate in the data? What is the median turnout rate? You may have to look up the functions for this. 

```{r}

```

2. Create a new variable named anes_vep_diff that computes the difference between the actual turnout rate (VEP_turnout) and ANES estimates of the turnout rate for each year. 

```{r}
# Notice that the ANES_turnout variable is on a different scale, where 50= 50%. So we need to multiply the previous variable we created by 100

```

3. How large is the difference on average? 

```{r}

```

4. What does this suggest about the prevalence of social desirability bias when it comes to reporting electoral behavior?




# Prepare for submission

You will submit all assignments and problem sets as Markdown files. To create a markdown file, click the Render icon. 


# Check-in

1.  On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class:
2.  What was the easiest thing to understand?
3.  What was the most difficult thing to understand?

# Before next class:

Review this notebook!

# Optional Resources

Canvas > Files > "Data Analysis Chapter 1" - Covers using R and basic operations


