---
title: "Class 15"
format: docx
execute:
  error: TRUE
---

# Plan
* Questions on group psets
* Uncertainty (Lecture)
* Standard error
* Group work: standard error for a difference in means
* Central Limit Theorem

# Uncertainty (lecture)

# Recap: how standard error works

Let's use the normal distribution to calculate the standard error of a poll of size 500.

First, let's simulate doing this manually:

```{r}

# You do not need to understand this code

population <- c(rep(1,50000),rep(0,50000))

# Simulation - Get sampling distribution
n <- 10000  # number of polls
ss <- 1000  # sample size
results <- rep(NA,n)
for (i in 1:n){
  poll <- sample(population,ss,replace=F)
  results[i] <- mean(poll)
}

# Sampling distribution
hist(results,breaks=50)

# Standard deviation of sampling distribution
round(sd(results),2)
```

Now let's use the formula (with only 1 poll)

```{r}
poll <- sample(population,ss,replace=F)
p_sd <- sd(poll)
se <- p_sd / sqrt(ss)
round(se, 2)
```


# Standard Error Beyond Polls

```{r}
library(tidyverse)
resume <- read.csv("./class_data/resume.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/resume.csv","resume.csv")
#resume <- read.csv("resume.csv")

```

Let's calculate the average callback rate in the data:

```{r}

```

Now we calculate the standard error:

```{r}

```

What's going on under the hood?

```{r}

# Simulation - Get sampling distribution
n <- 5000  # number of samples
results_resume <- rep(NA,n)
for (i in 1:n){
  resume_n <- resume %>% sample_n(4870, replace = TRUE)
  results_resume[i] <- mean(resume_n$call)
}

# Sampling distribution
hist(results_resume)

# Standard deviation of sampling distribution
sd(results_resume)

# Formula
sd(resume$call)/sqrt(nrow(resume))

```

One thing to keep in mind is that as we narrow the sample, the standard error will change (because the n and the sd change):

Let's get the standard error for the callback rate for white resumes.

```{r}
resume |> filter(race=="white") |> summarize()
```


# Group Work: Standard error for a difference in means

Let's work with the Chechen artillery fire natural experiment. First, obtain the difference in means between the treated and control group for the _postattack_ variable.

Treated group: (fire: 1)
Control group: (fire: 0)

```{r}
library(tidyverse)
chechen <- read.csv("./class_data/chechen.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/chechen.csv","chechen.csv")
#chechen <- read.csv("chechen.csv")

# Reminder: we can use group_by |> summarize() for this

```

Estimate the standard error for this difference in means, by obtaining the standard deviation and sample size of _each_ distribution. Remember that you can use table() to count observations; you can also use n() inside summarize() after doing a group_by(). To obtain the standard deviations for each group, use sd(outcome) on a filtered dataset, or use sd(outcome) within summarize()

```{r}


# Sample Size
n_treatment <- 
n_control <-  
  
# Standard deviation
sd_treatment <-
sd_control <-  

# Combine using formula
sqrt(((sd_treatment^2)/n_treatment) + ((sd_control^2)/ n_control) )
  
# Answer should match: 0.4165658
```

# Central Limit Theorem (lecture)

# Central Limit Theorem Demonstration

Plot the state population from earlier:

```{r}
hist(population)
```

Plot the means of the polls from earlier (the sampling distribution)

```{r}
# You do not need to know this code

hist(results,xlim=c(.4,.6),col="lightgray",breaks=(n/2))
abline(v=mean(results),col="blue",lwd=3)
abline(v=.5,col="red",lwd=3)

```

We have n values: how many are outside the boundaries? 

If it's a perfectly normal distribution, it should be ~2.5% on each side.

```{r}
# You do not need to know this code

length(results[results < mean(results) - 1.96*sd(results)])  / n
length(results[results > mean(results) + 1.96*sd(results)])  / n
```


# Confidence Interval (only if time)

Let's make a 90% confidence interval for our Chechnya estimate:

```{r}

```



## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
