---
title: "Class 5"
format: docx
execute:
  error: TRUE
---

# Plan
* Announcements
* Quiz Review
* Review: Group Work
* Natural experiments (brief lecture) 
* Group work - Chechnya natural experiment
* Distributions in R
* Group work 2

# Quiz Review

```{r}
fp <-  read.csv("class_data/foreignpopulation.csv")

# Non-git
#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/foreignpopulation.csv","foreignpopulation.csv")
#fp <- read.csv("foreignpopulation.csv")

# Get the average populations of countries in 2015 and 2011
fp |> summarize(mean(population_2015,na.rm=T)),mean(population_2011,na.rm=T))

```

# Review: Group Work

```{r}
library(tidyverse)

hsavings <- read.csv("./class_data/rosca.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/class_data/rosca.csv","rosca.csv")
#hsavings <- read.csv("rosca.csv")
```

Evaluate the quality of the randomization. To do so, test whether participants in each of the three groups (safe box, locked box, or control) have a similar age, on average. Then check whether a similar proportion of the three groups are female, and finally whether a similar proportion are married. 

```{r}

hsavings |> group_by(tgroup) |> 
  summarize(mean(bg_b1_age),mean(bg_female),mean(bg_married))

```

Now calculate the average treatment effect for those assigned to the safe box and locked box conditions. Which was more effective in increasing health savings?

```{r}

hsavings |> group_by(tgroup) |> 
  summarize(mean(fol2_amtinvest,na.rm=T))

# Safe box
408.2150	- 257.8333	

# Locked box
307.8261	- 257.8333
```

Was the locked box treatment more effective among male or female study participants? Recalculate the average treatment effects for each subset of participants

```{r}

hsavings |> group_by(tgroup,bg_female) |>
  summarize(mean(fol2_amtinvest,na.rm=T))

```


```{r}
crime  <- read.csv("class_data/crime.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/crime.csv","crime.csv")
#crime <- read.csv("crime.csv")
```

Evaluate the quality of the randomization by looking at the race of the applicant, as well as the type of job they applied to.

```{r}
crime %>%
group_by(crimrec) %>%
summarize()

```

Do you think this study likely identifies a causal relationship or a correlation? Why or why not? Thinking about internal vs. external validity, what are some potential limitations of this study?


# Natural experiments (brief lecture)

# Group work 1: Natural Experiment: Indiscriminate Violence and Insurgency

What is the relationship between state violence and insurgency? While some researchers and military analysts argue that indiscriminate violence is effective in suppressing insurgencies, many argue that violence actually increases insurgent attacks by creating cooperative relationships between affected citizens and insurgents. In this exercise, we will examine this question using data on indiscriminate Russian artillery fire in Chechnya from 2000 to 2005. 

```{r}
chechen <- read.csv("./class_data/chechen.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/chechen.csv","chechen.csv")
#chechen <- read.csv("chechen.csv")

```

The variables of interest are as follows:

fire            Whether Russians struck a village with artillery fire (equal to 1) or not (equal to 0)
deaths          Estimated number of individuals killed during Russian artillery fire
preattack       The number of insurgent attacks before Russian artillery fire
postattack      The number of insurgent attacks after Russian artillery fire

1. What proportion of villages were attacked by Russian artillery? What was the average number of deaths associated with an attack?

```{r}

```

2. Using the _postattack_ variable, compare the average number of insurgent attacks observed for villages hit by artillery fire vs. those that were not hit. 

```{r}

```

3. The results to the previous question should suggest that villages which experienced attacks had less insurgent activity afterwards. However, one possible explanation is that these villages also had less insurgent activity to begin with. This is possible because this is not an RCT. To examine this, look at the _preattack_ variable, which measures the time period before the Russian artillery attacks occurred. What is the difference in the average number of insurgent attacks between the villages that were later hit by artillery fire and those that were not? 

```{r}

```

4. Estimate the average treatment effect of indiscriminate fire on insurgency. The outcome variable is _postattack_

```{r}

```

5.Does this analysis support the claim that indiscriminate violence reduced insurgency attacks in Chechnya? If you were in charge of briefing military leaders on these results, what caveats and/or additional data might you want to draw on?



# Data Types and Distributions (Lecture)

To map a discrete variable to a new binary variable, we have to use logical statements.

This can be done using ifelse(logical condition, value if true, value if false). Let's create a new variable called control that receives a 1 if the person was in the control group, and 0 otherwise.

```{r}

hsavings <- hsavings |> mutate(control=ifelse(tgroup=="Control",1,0))

```

Now we can easily get the proportion of people assigned to the control group

```{r}


```

# Distributions in R

A histogram is useful for examining a single distribution. We use hist() to generate a histogram of a continuous variable. 

```{r}

hist(hsavings$bg_b1_age)
#abline(v=mean(,na.rm=T),col="red")

```

A boxplot is useful for comparing distributions. We use the formula boxplot(y~x), where y is the outcome variable and x is the group.

```{r}

boxplot( ~ )

```

We can find percentiles for a distribution by using quantile():

```{r}

quantile()

```


### Group Work 2: Distributions

1. Let's first change our outcome variable in the Chechen case to measure the change in insurgent attacks. This fixes any remaining imbalance that might exist in the preattack variable. To do so, create a new variable called diffattack that measures the change in attacks over time.

```{r}

chechen$diffattack <- chechen$postattack - chechen$preattack

```

2. Separately for the treated and control groups, calculate the median of diffattack as well as the 90th and 99th percentiles. Note: the quantile function can go in summarize()

```{r}


```

3. Create a histogram of diffattack

```{r}

```


4. Create a boxplot of diffattack, comparing the treatment and control groups:

```{r}



```


5. Average treatment effects can sometimes be driven by outliers. In other words, even if most units did not respond to the treatment, a large response among a small subset of treated units could generate an average difference between treated and control groups. For instance, consider the case where there are 10 villages. 9 villages did not change their resistance activity following artillery fire. However, the remaining village had a decrease of 10 attacks in the period following artillery fire. In this case the average treatment effect on diffattack would have been -1, even though most villages did not change their behavior. When we look at just the average treatment effect, we cannot distinguish this case from a hypothetical situation in which each of the 10 villages had 1 less attack following the artillery fire -- the average treatment effect would be identical (-1).

Based on your answers above, is the average treatment effect in the Chechen study likely influenced by outliers? In other words, is the observed reduction in violence concentrated in a subset of villages, or did most villages decrease violent resistance following artillery fire?

*Answer here*


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?