---
title: "Class 4"
format: docx
execute:
  error: TRUE
---

# Plan
* Lecture: RCTs
* Analyzing randomized experiments (RCTs) in R
* Break
* Quiz

# Brief Lecture 

## Analyze randomized treatments - RCTs

Gerber, Green & Latimer (2008) conducted an experiment which aimed to increase voter turnout in elections by exploiting social desirability bias. 

The research design randomly allocated citizens into one of four conditions:

* Neighbor Shaming - These individuals received a letter saying their neighbors would be told if they did not vote
* Hawthorne - These individuals received a letter saying researchers were observing whether they voted or not
* Civic Duty - These individuals received a letter saying it was a civic duty to vote
* Control - These individuals received no letter. 
  
After sending the letters, the researchers observed whether individuals voted in the 2008 primaries. 

Let's load the data 

```{r}
library(tidyverse)
social <- read.csv("class_data/social.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/social.csv","social.csv")
#social <- read.csv("social.csv")
```

The variables in the dataset are as follows:

hhsize          household size of voter
messages        GOTV messages voter received (Civic Duty, Control, Neighbors, Hawthorne)
sex             sex of voter (female or male)
yearofbirth     year of birth of voter
primary2008     whether a voter turned out in the 2008 Primary election (1=voted, 0=abstained)


Calculate the average turnout in the 2008 primaries for the 4 different groups ("Control","Civic Duty","Neighbors","Hawthorne"). 

```{r}

```

Calculate an average treatment effect for each treatment condition

```{r}

```

Evaluate randomization: For each of the 4 groups, what is the average household size? The average age in 2008? The proportion of females?

```{r}

```


# Group work 1: health savings experiment (please complete before turning in)

To understand why the poor are constrained in their ability to save for investments in preventative health products, researchers designed a field experiment in rural Kenya in which they randomly varied access to innovative saving technologies.  In the first treatment group (*Safe Box*), respondents were given a box locked with a padlock. This treatment is designed to estimate the effect of having a safe and designated storage technology for preventative health savings.  

In the second treatment group (*Locked Box*), respondents were given a locked box, but not the key to the padlock. The respondents were instructed to call the program officer once they had reached their saving goal, and the program officer would then meet the participant and open the box. Compared to the safe box, the locked box offered stronger commitment through earmarking (the money saved could only be used for health expenses). 

In this exercise, our outcome of interest is the amount (in Kenyan shilling) spent on preventative health products after 12 months.

| Name           | Description                                                     |
|----------------|-----------------------------------------------------------------|
| bg_female      | 1 if female, and 0 otherwise                                    |
| bg_married     | 1 if married, and 0 otherwise                                   |
| bg_b1_age      | age at baseline                                                 |
| tgroup         | Locked Box, Safe Box, or Control                                |
| fol2_amtinvest | Amount invested in health products (outcome)                    |
 
## Load the data

```{r}
hsavings <- read.csv("./class_data/rosca.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/rosca.csv","rosca.csv")
#hsavings <- read.csv("rosca.csv")
```

1. Evaluate the quality of the randomization. To do so, test whether participants in each of the three groups (safe box, locked box, or control) have a similar age, on average. Then check whether a similar proportion of the three groups are female, and finally whether a similar proportion are married. 

```{r}

```

2. Now calculate the average treatment effect for those assigned to the safe box and locked box conditions. Which was more effective in increasing health savings?

```{r}

```

3. Was the locked box treatment more effective among male or female study participants? Recalculate the average treatment effects for each subset of participants by filtering the data before calculating the treatment effect.

```{r}

```

## Group work 2: The effect of a criminal record (please complete before turning in)

In this exercise, we will analyze the effect of having a criminal record on one’s job prospects.

To examine the role of a criminal record, a researcher hired a pair of 23 year old white men and a pair of 23 year old black men and instructed them to apply for existing entry-level jobs in the city of Milwaukee. The men in each pair were matched on a number of dimensions, including physical appearance and self-presentation. As much as possible, the only difference between the two was that the researcher randomly varied which individual in the pair would indicate to potential employers that he had a criminal record related to drug offenses. Further, each week, the pair alternated which applicant would present himself as an ex-felon. In other words, each individual presented themselves as a felon half of the time. The researcher then compared callback rates.

Load the data:

```{r}
crime  <- read.csv("class_data/crime.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-f25/refs/heads/main/class_data/crime.csv","crime.csv")
#crime <- read.csv("crime.csv")

```

The relevant variables in the dataset are as follows:

call        1 if applicant received a callback, 0 if the applicant did not receive a callback
black       1 if the applicant identifies as black, 0 if the applicant does not identify as black
crimrec     1 if the applicant has a putative criminal record, 0 if the applicant does not (the treatment)
custserv    1 if job is in the costumer service sector, 0 if it is not
manualskill 1 if job requires manual skills, 0 if it does not
distance    Distance between the job and downtown Milwaukee

1. Calculate the average treatment effect of having a criminal record on the callback rate.

```{r}

```

2. Evaluate the quality of the randomization by looking at the race of the applicant, as well as the type of job they applied to.

```{r}

```

3. Do you think this study likely identifies a causal relationship or a correlation? Why or why not? Thinking about internal vs. external validity, what are some potential limitations of this study?

*Answer here*

## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?